---
title: "Functions and Iterations"
author: "Meenakshi Kushwaha"
date: "2022-08-18"

output: 
  html_document:
    toc: true
    toc_float: true
    code_download: true
    theme: flatly
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, cache = TRUE, warning = FALSE)
library(gapminder)
library(tidyverse)
library(here)
```

## Writing functions
In the `gapminder` dataset calculate the range of population, gdp, and life expectancy

```{r}
round(max(gapminder$lifeExp) - min(gapminder$lifeExp),1)

round(max(gapminder$pop) - min(gapminder$pop),1)

round(max(gapminder$pop) - min(gapminder$pop),1)
```

Make a function for calculating range
```{r}
max_minus_min <- function(x){
  round((max(x) - min(x)), 1)
  }
```

Test-run your function
```{r}
max_minus_min(gapminder$lifeExp)
```

## Iterations using `purrr` package

Apply `class()` function to each column of gampminder data

```{r}
gapminder %>% 
  map(class)

```

The default output of `map()` is a list. If you want a character vector as ouput use `map_chr()`

```{r}
gapminder %>% 
  map_chr(class)
```

Another example, What is the number of distinct values in each column? 
Hint: use `n_distinct()`

```{r}
gapminder %>% 
  map_int(n_distinct) # using map_int for integer output
```

What is the median of all numeric columns?
```{r}
gapminder %>% 
  dplyr::select_if(is.numeric) %>% 
  map_dbl(median)
```

## Using `~` with `map()`

## Reading multiple files using `purrr()`
Download the data here - [Google drive link](https://drive.google.com/drive/folders/1X54HsJUWnx9kAKYoTdceePoEP1139bKN?usp=sharing)

This is data from diffrent countries. "_gm" suffix is for gapminder from where data is borrowed. In the example below, the files in a folder called "data"


Step 1 and 2: Make a list of all `.csv` files with _gm suffix
```{r}
my_files <- dir(here("data"), # specify file path
                pattern = "*_gm.csv",  # look for .csv files with _gm suffix
                full.names = TRUE)  # preserve file path
my_files
```

Step 3: Read and combine all files using `map_dfr()`
```{r}

my_df <- my_files %>% 
  map_dfr(read_csv) 

my_df
```

