---
title: "Statistics and Modelling"
author: "Meenakshi Kushwaha"
date: "2022-09-01"

output: 
  html_document:
    toc: true
    toc_float: true
    code_download: true
    theme: flatly
    code_foalding: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, cache = TRUE, message = FALSE, warning = FALSE)
library(gapminder)
library(tidyverse)
library(here)
library(gtsummary)
library(flextable)
library(broom)
```

# Presenting Summary Statistics
## Simple table in R

```{r}
gapminder %>% 
  filter(country == "India",
         year > 1980) %>% 
  select(year, pop, lifeExp ) %>% 
  knitr::kable() # print in table format
```

## Summary statistics
Generate summary statistics by continent for `gapminder` dataset

```{r}
gapminder %>% 
  select(-country) %>% # drop country and year variable
  tbl_summary()
```

## Customizing your table

### Include/exclude variables
```{r}
gapminder %>% 
  select(-country) %>% # drop country variable
  tbl_summary(include = -year) # include everything but year
```


```{r}
gapminder %>% 
  select(-country) %>% # drop country variable
  tbl_summary(include = c(continent, lifeExp, gdpPercap))
```

Format labels
```{r}
gapminder %>% 
  select(-country, - year) %>% # drop country and year variable
  tbl_summary() %>% 
  bold_labels()
```

### Saving your table
Save your table as .docx file
```{r}
gapminder %>% 
  select(-country, - year) %>% # drop country and year variable
  tbl_summary() %>% 
  as_flex_table() %>% # save table as Word file 
  save_as_docx(path = here("myfile.docx")) 
```

## Simple linear models in R

```{r}
lm(lifeExp ~ pop, data = gapminder)
```

Store model as object 
```{r}
my_model <- gapminder %>% 
  lm(lifeExp ~ year, data = .) 
summary(my_model)
```

## Testing model assumptions

Make a model for Argentina and plot diagnostics
```{r}
arg_model <- gapminder %>% 
  filter(country == "Argentina") %>% 
  lm(lifeExp ~ year, data = .)
plot(arg_model) # generates four diagnostic plots
```

Try the same for Rwanda and compare
```{r}
rwanda_model <- gapminder %>% 
  filter(country == "Rwanda") %>% 
  lm(lifeExp ~ year, data = .)
plot(rwanda_model) # generates four diagnostic plots
```

## Multiple models [Home Exercise] 
How does life expectancy change with time in each country?

###  Global vs local trend
```{r}
gapminder %>%
  ggplot(aes(x = year, y = lifeExp, group = country)) +
  geom_line(alpha = 1/3)
```


### Keep related things together `group_nest()`
Nest by country
```{r}
by_country <- gapminder %>% 
  mutate(year1950 = year - 1950) %>% 
  group_nest(continent, country)
head(by_country)
```

### Linear model per country
```{r}
by_country_lm <- by_country %>% 
  mutate(model = map(data, ~lm(lifeExp ~ year1950, data = .)))# to each element of data column apply lm() function
by_country_lm
```

### Plot for one country
```{r}
by_country_lm %>% 
  filter(country=="India") %>% 
  unnest(data) %>% 
  ggplot(aes(x = year1950, y = lifeExp)) +
  geom_line()
```

### Linear model for India
```{r}
by_country_lm%>% 
  filter(country == "India") %>% 
  pluck("model", 1) %>% 
  summary()
```

### Linear model for Rwanda
```{r}
by_country_lm%>% 
  filter(country == "Rwanda") %>% 
  pluck("model", 1) %>% 
  summary()
```

### Verify model results by plotting for Rwanda
```{r}
by_country_lm %>% 
  filter(country=="Rwanda") %>% 
  unnest(data) %>% 
  ggplot(aes(x = year1950, y = lifeExp)) +
  geom_line()
```

### Cleaning model results with `broom`

```{r}
gm_models <- by_country_lm %>% 
  mutate(glance= map(model, glance),
         tidy = map(model, tidy),
         augment = map(model, augment),
         rsq = map_dbl(glance, "r.squared"))
gm_models
```

### Visualizing all models
```{r}
gm_models %>%
  unnest(tidy) %>%
  select(continent, country, rsq, term, estimate) %>%
  pivot_wider(names_from = term, values_from = estimate) %>%
  ggplot(aes(x = `(Intercept)`, y = year1950)) +
  geom_point(aes(colour = continent, size = rsq)) +
  geom_smooth(se = FALSE, method = "loess") +
  scale_size_area() + labs(x = "Life expectancy (1950)", y = "Yearly improvement")