---
title: "Data Wrangling"
author: "Meenakshi Kushwaha"
date: '2022-07-21'
output: 
  html_document:
    toc: true
    toc_float: true
    code_download: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      warning = FALSE, 
                      cache = TRUE, 
                      message = FALSE)
```

## Load packages
```{r message=FALSE, warning=FALSE}
library(tidyverse) # data wrangling
library(gapminder) # dataset
```

Take a look at the dataset. **Tibbles** are a type of dataframes that work better with tidyverse. To see the full dataset use View()
```{r dataset}
gapminder # print the dataset
```
## Filter rows or observations using `filter()`
```{r filter}
gapminder %>% 
  filter(country == "India") # filter observations according to specific country

gapminder %>% 
  filter(country == "India", year < 1975) # filter using two conditions

gapminder %>% 
  filter(country=="India" | country == "Nepal") # filter using "or" condition

gapminder %>% 
  filter(year %in% c(1952,1962, 1972 )) # filter using multiple values

```

## Arrange rows or observations using `arrange()`

```{r arrange}
gapminder %>% 
  filter(year %in% c(1952,1962, 1972 )) %>% 
  arrange(year) # arrange in ascending order by default

gapminder %>% 
  filter(year %in% c(1952,1962, 1972 )) %>% 
  arrange(desc(year)) # arrange in descending order

```


## Select variables of interest using `select()`

```{r select}
gapminder %>% 
  select(country, year, pop) # select variables of interest

gapminder %>% 
  select(-pop) # drop variables using -
```

## Create new variable using `mutate()`
```{r mutate}
gapminder %>% 
  select(country, pop) %>% 
  mutate(pop_mil = round(pop/1000000, 1)) # add a new variable

gapminder %>% 
  select(country, year, pop) %>% 
  mutate(pop_rank = min_rank(pop)) %>%  # add new variable using rank
  arrange(pop_rank) # arrange in ascening order 

gapminder %>% 
  select(country, pop) %>% 
  transmute(pop_mil = round(pop/1000000, 1)) # only keep the new variable 
```

### Creating new variable based on some condition(s)
```{r case_when}
gapminder %>% 
  select (country, continent, year, pop) %>% 
  mutate(cont_code = case_when(continent == "Africa" ~ 1,
                               continent == "Americas" ~ 2,
                               continent == "Asia" ~ 3,
                               continent == "Europe" ~ 4,
                               TRUE ~ 5)) ## option when none of the above conditions are true
```

## Summarise your dataset 
```{r summarise}
gapminder %>% 
  summarise(mean_lifeExp = mean(lifeExp)) # summarise the entire data in one line

gapminder %>% 
  group_by(continent) %>% # group by continent
  summarise(mean_lifeExp = mean(lifeExp)) # summarise by groups
```

## Count number of observations in each group
```{r counts}
# both codes below will give the same output

gapminder %>% 
  count(continent)

gapminder %>% 
  group_by(continent) %>% 
  summarise(observations = n())


```

## Exercises
### Filter the data for India and Nepal for the years 1997, 2002, and 2007
```{r exercise}
gapminder %>% 
  filter(country == "India"| country == "Nepal",
         year %in% c(1997, 2002, 2007))
```
### What is the mean life expectancy of countries in Asia

```{r}
gapminder %>% 
  filter(continent == "Asia") %>% 
  group_by(country) %>% 
  summarise(mean(lifeExp))
```

### In 2007 Which countries have the highest and lowest gdppercap in Asia

```{r}
gapminder %>% 
  filter(continent == "Asia", year == 2007) %>% 
  group_by(country) %>% 
  summarise(avg_gdp = mean(gdpPercap)) %>% 
  arrange(avg_gdp) %>% 
  tail()
```

